Morus: Biases in the stream

### 2018-05-08 morusAC_02_Preliminaries.tex

Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
\begin{align*}
\textrm{Round 1:} &&
S^{t+1}_0 &\leftarrow ( S_t^0 \oplus S^t_1 \cdot S^t_2 \oplus S^t_3 ) \lll_w b_0 &
S^t_3 \leftarrow S^t_3 \lll b'_0.\\
\textrm{Round 2:} &&
S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
S^t_4 \leftarrow S^t_3 \lll b'_1.\\
\textrm{Round 3:} &&
S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
S^t_0 \leftarrow S^t_3 \lll b'_2.\\
\textrm{Round 4:} &&
S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
S^t_1 \leftarrow S^t_3 \lll b'_3.\\
\textrm{Round 5:} &&
S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
S^t_2 \leftarrow S^t_3 \lll b'_4.
\end{align*}

\subsubsection{Initialization.}
The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K$ and a 128-bit nonce IV into the state, and running the {\tt StateUpdate} for 16 steps. The key and IV are loaded into the state as follows:
\begin{align*}
S^{-16}_0 = \textrm{IV}_{128}, &&
S^{-16}_1 = K_{128}, &&
S^{-16}_2 = 1^{128}, &&
S^{-16}_3 = const_0, &&
S^{-16}_4 = const_1.
\end{align*}
Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.

The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size. The state is initialized as
\begin{align*}
S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
S^{-16}_1 &= K_{128}\|K_{128}, &
S^{-16}_2 = 1^{256}, \\
S^{-16}_3 &= 0^{256}, &
S^{-16}_4 &= const_1\|const_0.
\end{align*}
After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.

>>>>

Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
\begin{align*}
\textrm{Round 1:} &&
S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 ) \lll_w b_0 &
S^t_3 \leftarrow S^t_3 \lll b'_0.\\
\textrm{Round 2:} &&
S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
S^t_4 \leftarrow S^t_4 \lll b'_1.\\
\textrm{Round 3:} &&
S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
S^t_0 \leftarrow S^t_0 \lll b'_2.\\
\textrm{Round 4:} &&
S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
S^t_1 \leftarrow S^t_1 \lll b'_3.\\
\textrm{Round 5:} &&
S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
S^t_2 \leftarrow S^t_2 \lll b'_4.
\end{align*}

\subsubsection{Initialization.}
The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running the {\tt StateUpdate} for 16 steps. The key and nonce are loaded into the state as follows:
\begin{align*}
S^{-16}_0 = \textrm{IV}_{128}, &&
S^{-16}_1 = K_{128}, &&
S^{-16}_2 = 1^{128}, &&
S^{-16}_3 = const_0, &&
S^{-16}_4 = const_1.
\end{align*}
Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.

The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size and the two possible key size:
\begin{itemize}
  \item[-] When the key size is 128 bits, $K = K_{128} \|K_{128}$
  \item[-] When the key size is 256 bits, $K = K_{256}$
\end{itemize}
The state is initialized as
\begin{align*}
S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
S^{-16}_1 &= K, &
S^{-16}_2 = 1^{256}, \\
S^{-16}_3 &= 0^{256}, &
S^{-16}_4 &= const_1\|const_0.
\end{align*}
After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K$.
